---
title: "coding_challenge_7"
author: "Radhika"
date: "2026-03-26"
output: 
  html_document:
  md_document: 
  variannt: gfm 
---
##1.Read in the data called “PlantEmergence.csv” using a relative file path and load thefollowing libraries. tidyverse, lme4, emmeans, multcomp, and multcompView. Turn the Treatment , DaysAfterPlanting and Rep into factors using the function as.factor

```{r 1. }

plant<- read.csv("PlantEmergence.csv")
library(tidyverse)
library(lme4)
library(emmeans)
library(multcomp)
library(multcompView)
plant 

plant$Treatment<- as.factor(plant$Treatment)
str(plant$Treatment)

plant$DaysAfterPlanting<- as.factor(plant$DaysAfterPlanting)

plant$Rep<- as.factor(plant$Rep)
str(plant)

```

##2. Fit a linear model to predict Emergence using Treatment and DaysAfterPlanting along with the interaction. Provide the summary of the linear model and ANOVA results.

```{r 2.}

model1<-lm(Emergence~Treatment*DaysAfterPlanting, data= plant)
summary(lm(Emergence~Treatment*DaysAfterPlanting, data= plant))

anova(lm(Emergence~Treatment*DaysAfterPlanting, data= plant))


```

##3. Based on the results of the linear model in question 2, do you need to fit the interaction term? Provide a simplified linear model without the interaction term but still testing both main effects. Provide the summary and ANOVA results. Then, interpret the intercept and the coefficient for Treatment 2.

The interaction is not signifcant strategically. so we don't need to include that in our model. 


```{r 3. }

model2<-lm(Emergence~Treatment+DaysAfterPlanting, data= plant)
summary(lm(Emergence~Treatment+DaysAfterPlanting, data= plant))

anova(model2)

```
the intercept for model2 is   182.163 and coefficient for treatment 2 is  -134.531. it is negative showing negative relationship between the two variables. treatment2 has a significant effect on the dependent variable. 

##4. Calculate the least square means for Treatment using the emmeans package and perform a Tukey separation with the compact letter display using the cld function. Interpret the results.

```{r 4. }
lsmeans<- emmeans(model2, ~Treatment )
Results_lsmeans <- cld(lsmeans)

Results_lsmeans 
```
Treatment 2 has the lowest mean and treatment 7 has the intermediate mean while treatment 6 has highest mean. there are different groups, groups with common number are not significant different from each other. so treatment 2 is significant diffrent from all other treatment groups. treatment 7 is different from all except 1, 9, 4, 8. 6, 3, 5 treatment are significantly different from 2 and 7. 


##5. he provided function lets you dynamically add a linear model plus one factor from that model and plots a bar chart with letters denoting treatment differences. Use this model to generate the plot shown below. Explain the significance of the letters.

```{r 5. }
plot_cldbars_onefactor <- function(lm_model, factor) {
data <- lm_model$model
variables <- colnames(lm_model$model)
dependent_var <- variables[1]
independent_var <- variables[2:length(variables)]
lsmeans <- emmeans(lm_model, as.formula(paste("~", factor))) # estimate
lsmeans
Results_lsmeans <- cld(lsmeans, alpha = 0.05, reversed = TRUE, details =
TRUE, Letters = letters) # contrast with Tukey adjustment by default.
# Extracting the letters for the bars
sig.diff.letters <- data.frame(Results_lsmeans$emmeans[,1],
str_trim(Results_lsmeans$emmeans[,7]))
colnames(sig.diff.letters) <- c(factor, "Letters")
# for plotting with letters from significance test
ave_stand2 <- lm_model$model %>%
group_by(!!sym(factor)) %>%
dplyr::summarize(
ave.emerge = mean(.data[[dependent_var]], na.rm = TRUE),
se = sd(.data[[dependent_var]]) / sqrt(n())
) %>%
left_join(sig.diff.letters, by = factor) %>%
mutate(letter_position = ave.emerge + 10 * se)
plot <- ggplot(data, aes(x = !! sym(factor), y = !! sym(dependent_var))) +
stat_summary(fun = mean, geom = "bar") +
stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.5) +
ylab("Number of emerged plants") +
geom_jitter(width = 0.02, alpha = 0.5) +
geom_text(data = ave_stand2, aes(label = Letters, y = letter_position),
size = 5) +
xlab(as.character(factor)) +
theme_classic()
return(plot)
}

plot_cldbars_onefactor(model2, "Treatment")
```


Treatment 2 is showing statistically significant difference from rest of the treatments. Treatment 7 does not show statistically significant diffrence from 1,4,8 ,9 treatments but shows statistically significant difference from treatments 2,3,5, 6. 
At treatment 2, the number of emerged plants are singnificantly lower than any other treatment. treatment 7 is showing intermediate number of emerged plants along with treatments 1,4,8 and 9 treatments. 
