---
title: "2_april_classnotes"
author: "Radhika"
date: "2026-04-02"
output:
  
  html_document:
  md_document:
    variant: gfm 
---
# Introduction

Packages are installed in system library. all of the functions, code are stored in our computer, which can leads to reproduibility problem.

-why we need this?
Packages can change over time. 

lock file contains all the packages which will be uploaded into github. all the packages that we used can be installed by anyone using our github. 


# renv

1. install r env
2. renv::init(). this code is used in console and run.
3. all the packages information will be displayed in console. 
4. it has created a directory, which contains library folder having all functions and codes wrt the packages. 

source("renv/activate.R")

.libPaths() when run in console will shows where the project is pulling the r packages from. 
"/Library/Frameworks/R.framework/Versions/4.5-arm64/Resources/library" is shows where the packages will be pulled from. 

"it basically encodes the same version u are using, that the end user will use"

# when we want to update the packages

install.packages("MASS") any package. 
go to files, lock file has all the pacakges. 
```{r}

library(ggplot2)
library(dplyr)
#library(paletteer)

```


renv::snapshot - like save button
it will ask to proceed?- yes to upload the new packages.


# when third person downloads the zip file from github. 

he needs to use restore function to get the packages. 

-renv::restore()
-do u want to proceed?- yes
 its' gonna look into ur computer first. if the package does not exist, it will install them. 
 
 
# Limitations

1. pandock , its itegrate with r markdown. its a outside package. so the person first needs to install pandock to use it. 


-------------SECOND PART------------------------------------------------------------------------


```{r}

mtcars
#subsetting

mtcars[1, 4]

#finds all rows such that vs is 1
mtcars[mtcars$vs==1, ]

mtcars$hp[mtcars$vs==1 & mtcars$gear== 4]
#using subset function
subset(mtcars, vs==1 | gear==4)

#cars with high hp

max(mtcars$hp)

```
















