---
title: "coding_challenge_6"
author: "Radhika"
date: "2026-03-19"
output: 
  md_document: 
  variant: gfm
  html_document:
---

# 1.

writing your own functions and iterations helps in reducing the error that might occur during copy and pasting the code. it makes our work more reproducible and makes it easier for someone else to understand. 

# 2. 
  1. we just have to use the function(). define it by the name we wanted to use our function followed by {} in which we write the formula that will be used everytime the name of the defined function is used. return will help in returning the value.  
  
  2. for the loop, for() is used, in which we define the sequence on which we wanted to apply the code. it is followed by print where values are returned in the form of vector, list or dataframe. 
  
# 3., 4 and 5. 
```{r code}
cities<- read.csv("Cities.csv")

distance <- function(lat1, lon1, lat2, lon2){
  # convert to radians
rad.lat1 <- lat1 * pi/180
rad.lon1 <- lon1 * pi/180
rad.lat2 <- lat2 * pi/180
rad.lon2 <- lon2 * pi/180
# Haversine formula
delta_lat <- rad.lat2 - rad.lat1
delta_lon <- rad.lon2 - rad.lon1
a <- sin(delta_lat / 2)^2 + cos(rad.lat1) * cos(rad.lat2) * sin(delta_lon / 2)^2
c <- 2 * asin(sqrt(a))
# Earth's radius in kilometers
earth_radius <- 6378137
distance_km <- (earth_radius * c)/1000
return(distance_km)
}

# 5.  distance between new york city and auburn

auburn<- cities$lat[cities$city== "Auburn"]

                 

distance(
  lat1= cities$lat[cities$city== "Auburn"], 
  lon1 = cities$long[cities$city== "Auburn"], 
  lat2 = cities$lat[cities$city== "New York"],
  lon2 = cities$long[cities$city== "New York"]
)



```

# 6.

```{r one way}
all_cities<- distance(
  lat1= cities$lat[cities$city== "Auburn"], 
  lon1 = cities$long[cities$city== "Auburn"],
  lat2 = cities$lat,
  lon2 = cities$long
)

all_cities
```


```{r using loop function}
dataframe<- NULL
unique_cities = unique(cities$city)
for (i in seq_along(unique_cities)){
  lat_city = cities$lat[cities$city==unique_cities[[i]]]
  long_city = cities$long[cities$city == unique_cities[[i]]]
  distance_i <- distance(cities$lat[cities$city== "Auburn"], cities$long[cities$city== "Auburn"],lat_city, long_city)
  table <- data.frame(unique_cities[[i]], "Auburn", distance_i)
 dataframe<- rbind.data.frame(dataframe, table) 
}

```

 

```{r}
dataframe
colnames(dataframe) <- c("City1", "City2", "Distance_Km")
dataframe


```

