---
title: "CodingChallenge5"
author: "Radhika"
date: "2026-03-05"
output: html_document
md_document:
    variant: gfm
---

```{r reading the csv files}
df1<- read.csv("DiversityData.csv")
df2<- read.csv("Metadata.csv")
library(tidyverse)
#joining the two df by code 
alpha<- inner_join(df1, df2, by ="Code")

#3.Calculate Pielou’s evenness index: Pielou’s evenness is an ecological parameter calculated by
#the Shannon diversity index (column Shannon) divided by the log of the richness column.
#a. Using mutate, create a new column to calculate Pielou’s evenness index.
#b. Name the resulting dataframe alpha_even.

alpha_even<- mutate(alpha, even= shannon/log(richness))
alpha_even

#4. Using tidyverse language of functions and the pipe, use the summarise function and tell me
#the mean and standard error evenness grouped by crop over time.
library(tidyverse)
alpha_average<- alpha_even %>%
  group_by(Crop, Time_Point)%>%
  summarise(mean.even= mean(even), 
            n = n(), 
            sd.dev = sd(even))%>%
  mutate(std.err = sd.dev/sqrt(n))





```



Calculate the difference between the soybean column, the soil column, and the difference
between the cotton column and the soil column
a. Start with the alpha_average dataframe
b. Select relevant columns: select the columns Time_Point, Crop, and mean.even.
c. Reshape the data: Use the pivot_wider function to transform the data from long to wide
format, creating new columns for each Crop with values from mean.even.
d. Calculate differences: Create new columns named diff.cotton.even and diff.soybean.even
by calculating the difference between Soil and Cotton, and Soil and Soybean,
respectively.
e. Name the resulting dataframe alpha_average2


```{r 5th question}
alpha_average2<- alpha_average%>%
  select(Time_Point, Crop, mean.even)%>%
  pivot_wider(names_from = Crop, values_from = mean.even)%>%
  mutate(diff.cotton.even = Soil - Cotton)%>%
  mutate(diff.soybean.even = Soil - Soybean)
alpha_average2
```

Connecting it to plots
a. Start with the alpha_average2 dataframe
b. Select relevant columns: select the columns Time_Point, diff.cotton.even, and
diff.soybean.even.
c. Reshape the data: Use the pivot_longer function to transform the data from wide to long
format, creating a new column named diff that contains the values from diff.cotton.even
and diff.soybean.even.
 i. This might be challenging, so I’ll give you a break. The code is below.
pivot_longer(c(diff.cotton.even, diff.soybean.even),
names_to = "diff")
d. Create the plot: Use ggplot and geom_line() with ‘Time_Point’ on the x-axis, the column
‘values’ on the y-axis, and different colors for each ‘diff’ category. The column named
‘values’ come from the pivot_longer. The resulting plot should look like the one to the
right.

```{r 6th question }
alpha_average2 %>%
  select(Time_Point, diff.cotton.even, diff.soybean.even) %>%
  pivot_longer(c(diff.cotton.even, diff.soybean.even),
names_to = "diff")%>%
  ggplot( aes(x = Time_Point, y = value, color= diff) ) +
  geom_line()
```

